Species motif participation provides unique information about species risk of extinction

### 2023-03-31 manuscript/cover_letter.tex

\address{Anna {\AA}kesson \\ Link\"oping University \\ Division of Theoretical Biology, Dept.~IFM \\ SE-58183 Link\"oping, Sweden}

\date{October 21, 2021}

@@

Prof.~Barrett\\
Editor-in-Chief\\
Proceedings of the Royal Society B} % Name/title of the addressee

\opening{Dear Prof.~Barrett,}

We invite you to consider the enclosed ``Motif participation provides unique information about species risk of extinction'' by Anna {\AA}kesson et al., for publication in \emph{Proceedings of the Royal Society B} as an Article. Dr. Cirtwill was approached by the Proceedings B Preprint Editorial Team to consider publishing a thematically similar paper in this journal. That manuscript is currently in revisions elsewhere, but we believe that this manuscript also would be a good topical fit.

Ecological effects of disturbances in food webs are known to depend on global (e.g. food web connectance) and local (e.g. trophic level) network properties. Disturbances to the basal level of a food web have a particularly strong impact on the persistence of all species. 
However, the locations for secondary extinctions are difficult to identify as both direct and indirect interactions among species are important. As global network properties are too coarse-grained to include interactions, and local network properties disregard any indirect effects, we must consider other structural properties in order to fully understand secondary extinctions. Meso-scale network properties, such as smaller sub-graphs within a large network (motifs), offer a potentially valuable middle-ground approach. Motifs have been shown to be important for community-level stability, but how motifs relate to risk of secondary extinctions is so far unexplored.

@@

As we take an important step towards understanding of how secondary effects propagate in complex food webs, we believe our method and results will interest a wide audience of ecologists including both theoreticians and empiricists. We therefore think \emph{ Proceedings of the Royal Society B} would be an ideal platform for our research.

@@

\hspace{4\parskip} Anna {\AA}kesson, Alyssa R. Cirtwill, Kate Wootton, and Anna Ekl\"of

\vspace{4\parskip}
\centering
\includegraphics[width=60mm]{manuscript/figures/LiU_primary_black.pdf}

>>>>

\address{Alyssa Cirtwill \\ University of Helsinki \\ Spatial Foodweb Ecology Group \\ Helsinki, Finland}

\date{October XX, 2022}

@@

Editorial Board\\
Journal of Animal Ecology} % Name/title of the addressee

\opening{Dear Editor,}

We invite you to consider the enclosed manuscript ``Species motif participation provides unique information about species risk of extinction'' for publication in the \emph{Journal of Animal Ecology} as an Article.

Ecological effects of disturbances in food webs are known to depend on global (e.g. food web connectance) and local (e.g. trophic level) network properties. Disturbances to the basal level of a food web have a particularly strong impact on the persistence of all species. 
However, the locations for secondary extinctions are difficult to identify as both direct and indirect interactions among species are important. As global network properties are too coarse-grained to include interactions and local network properties disregard any indirect effects, we must consider other structural properties in order to fully understand secondary extinctions. Meso-scale network properties, such as smaller sub-graphs within a large network (motifs), offer a potentially valuable middle-ground approach. Motifs have been shown to be important for community-level stability, but how motifs relate to risk of secondary extinctions is so far unexplored.

@@

As we take an important step towards understanding of how secondary effects propagate in complex food webs, we believe our method and results will interest a wide audience of ecologists including both theoreticians and empiricists. We therefore think the \emph{Journal of Animal Ecology} would be an ideal platform for our research.

@@

\hspace{4\parskip} Alyssa R. Cirtwill, Anna {\AA}kesson,  Kate Wootton, and Anna Ekl\"of

\vspace{4\parskip}
\centering
\includegraphics[width=60mm]{HY_logo_Eng.eps}
